### 2011-05-09 submitting.tex
some initial tries
git-svn-id: https://svn.mathsim.eu/documents/publications/journal/style@2675 a653633e-0a08-0410-9ecb-ed4064e9718a

% $Id$
\documentclass{ansarticle}
%
%
\title{Preparing and submitting articles for the Archive of Numerical
  Software}
\author[1]{Wolfgang Bangerth}
\author[2]{Peter Bastian}
\author[1]{Guido Kanschat\thanks{NSF}}
\author[3]{Michael Heroux}
\author[4]{Anders Logg}
\affil[1]{Department of Mathematics, Texas A\&M University}
\affil[2]{IWR, Universit\"at Heidelberg}
\affil[3]{Sandia National Laboratories}
\affil[4]{Simula, Oslo}
\begin{document}
\maketitle
\begin{abstract}
  This document describes the general structure of the
  \texttt{ansarticle} document class and serves as an example for its
  usage. \texttt{ansarticle.cls} is based on the standard \LaTeX{}
  \texttt{article.cls} as well as a few other standard packages. It
  fixes most style parameters to the preferred style of the Archive of
  Numerical Software.
\end{abstract}
\section{Implementation}
The file \texttt{ansarticle.cls} loads the standard LaTeX
\texttt{article.cls} and applies some changes to the layout.
\subsection{Requirements}
Upon loading, \texttt{ansarticle.cls} requires the following packages:
\begin{itemize}
\item \texttt{times} and \texttt{mathtpm} to use postscript fonts
\item \texttt{hyperref} in order to typeset hyper-links in the text
\item \texttt{fancyhdr} for page headers and footers
\item \texttt{authblk} for more convenient handling of author and
  affiliation lists
\item \texttt{listings} for source code listings
\end{itemize}
These packages are part of standard \TeX{} distributions.
\section{Usage}
\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: